Flowchart of Current Algorithm

### 2009-05-04 docs/flowchart/flowchart.tex
lots of changes, primarily, switched the advect_base_pres stuff and put in the enforce_HSE method

\item For any quantity, $\phi,$ we define $\overline{\phi} = ${\bf Avg}$(\phi),$ 
the average over $\Omega_H,$ as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
here $\Omega_H$ is a layer at constant radius $r$.  In plane-parallel, this is 
as straightforward as averaging cell values at a particular height.  For spherical 
problems, see the description in Section \ref{Sec:Avg}.

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite VODE).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
from equation (8) in paper III,
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc)^{\outp} = -\sum_k(\rho\omegadot_k)^{\outp} \; q_k \enskip.
\end{equation}
for the case where only strong-mediated reactions are involved.  As
$\Hnuc$ is an output of the reaction network, more general networks
(i.e.\ involving weak interactions) can easily be added.  The enthalpy
update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
updated by
\begin{equation}
(\rho h)^{\outp} = 
(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$\\
is the process by which we update the base state density through $\dt$ in time
as well as compute time-centered interface states for use later in the algorithm.
\begin{description}

\item[plane-parallel:] We discretize the base state density evolution 
equation from paper III equation (29), without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are found via the
procedure described in paper II, Appendix A.\\

\item[spherical:] The base state density update now includes the area factors in the 
divergences:
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
          \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right] \enskip. 
\end{equation}
\end{description}
\item {\bf Advect Base Pressure}
is the process by which we update the base state pressure through $\dt$ in time and
compute $\psi$.  The input arguments are different for plane-parallel and spherical.
\begin{description}
\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
[p_0^\outp,\psi^\outp]$:

First we compute a cell-centered $\psi^\outp$ using
\begin{equation}
 \psi_j^\outp = \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
\end{equation}
Then, we discretize the base state pressure evolution equation from paper III equation (31),
\begin{equation}
\frac{\partial p_0}{\partial t} = 
-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip ,
\label{eq:p0upd_new}
\end{equation}
to compute the new base state pressure,
\begin{equation}
p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\outp 
\enskip,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.\\

\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp},\psi^{n+\myhalf,\outp}]$:

First use a predictor-corrector formulation for the pressure update 
in order to improve the overall accuracy.  Beginning with equation (\ref{eq:w01}),
we can write
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
 \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) - \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
\end{equation}
We time-center the left hand side:
\begin{equation}
\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
   \frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} - p_0^{\inp}}{\dt}\enskip.
\end{equation}
We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
\begin{equation}
p_0^{\star} = p_0^{\inp} \frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
\end{equation}
with:
\begin{equation}
f = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} \left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
\end{equation}
where the $(\partial p_0^\inp / \partial r)$ is computed by
interpolating $p_0^\inp$ from cell-centers to edges and then
differencing.  Next, we compute a provisional updated $\gammabar$ using the
provisional updated pressure,
$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
$X_k$ which were used to compute $\gammabar^\inp$.\\

Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
we can compute a more accurate approximation to equation (\ref{eq:frhs}),
\begin{equation}
f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -  \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
                                             \frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} \left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
\end{equation}
where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
This allows us to complete the corrector for the pressure update:
\begin{equation}
p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
\end{equation}
Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  First, we compute an updated $\gammabar$ using the update pressure, $\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.  Then, rewriting equation (\ref{eq:w01}), we can define
\begin{eqnarray}
\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
\end{eqnarray}
We discretize this locally as:
\begin{equation}
\psi_j^{\outp,n} = 
\left(\gammabar^\inp p_0^\inp\right)_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
\end{equation}
\begin{equation}
\psi_j^{\outp,n+\myhalf} = 
\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
\end{equation}
\end{description}

@@

We discretize the base state enthalpy equation,
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.\\

@@

The base state enthalpy update now includes the area factors in the divergences:
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left \{ \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j+\myhalf} -  
          \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j-\myhalf} \right\}
+\Delta t\psi^{\inp,n+\myhalf} \enskip. 
\end{equation}
The other difference is that the interface states are computed using 
$\psi^{\inp,n}$ as the forcing.

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to adjust the base state enthalpy
to account for large-scale mixing, reactions, and thermal diffusion.
Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
\begin{description}
\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
\end{description}

@@

\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
is the process by which we compute gravity at both radial cell-centers and edges.
\begin{description}
\item For plane-parallel, we hold the gravity constant in space and time.
\item For spherical, we define the enclosed mass at edges, and then cell-centers:
\begin{equation}
m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
\end{equation}
Then, we define the gravity at both edges and cell-centers:
\begin{equation}
g_{j-\myhalf} = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ g_j = \frac{Gm_j}{r_j^2}\enskip .
\end{equation}
\end{description}

@@

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3.] {\em React the full state and base state through the first time 
interval of $\dt / 2.$}

@@

\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n]
                   \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
                                  (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$

@@

\item Update $p_0$ and compute $\psi$.
\begin{description}
\item For plane-parallel,\\
{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
\item For spherical,\\
{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.

>>>>

\item For any quantity, $\phi$, we define $\overline{\phi} = ${\bf Avg}$(\phi)$
as the average over a layer at constant radius r, $\Omega_H$, as
\begin{equation}
\overline{\phi(r)}
= \frac{1}{\mathrm{A}(\Omega_H)}\int_{\Omega_H} \phi(r,{\bf x}) \; dA \enskip .
\end{equation}
In plane-parallel, this is as straightforward as averaging cell values at 
a particular height.  For spherical problems, see the description in 
paper IV, section 2.2 (or section \ref{Sec:Avg}).

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}, 
p_0^{\inp}]\rightarrow [\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, 
(\rho \omegadot_k)^{\outp}, (\rho\Hnuc)^{\outp}]$ \\
is the process by which we evolve the species and enthalpy from 
$X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following 
system of equations over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )
\enskip. \label{eq:reacttemp}
\end{eqnarray}
using the stiff ordinary differential equation integration methods
provided by the {\tt VODE} package (cite vode).  The absolute error
tolerances are set to $10^{-12}$ for the species, and a relative
tolerance of $10^{-5}$ is used for the temperature.  The integration
yields the new values of the mass fractions, $X_k^{\outp}$.  Equation
(\ref{eq:reacttemp}) is derived from the full temperature evolution equation
from equation (8) in paper III,
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
by assuming that the pressure is constant during the burn state.  In
evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  In
theory, this means evaluating the equation of state for each
right-hand side evaluation that {\tt VODE} requires.  In practice, we
freeze $c_p$ and $\xi_k$ at the start of the integration time step and
compute them using $\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs
to the equation of state.  We update temperature using
$T^{\outp} = T(\rho^\outp,h^\outp,X_k^\outp)$ if {\tt use\_tfromp = FALSE} or 
$T^{\outp} = T(\rho^\outp,p_0,X_k^\outp)$  if {\tt use\_tfromp = TRUE}.  
Note that the density remains unchanged
during the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

After the new mass fractions have been computed, the reaction rates are defined as:
\begin{equation}
(\rho\omegadot_k)^{\outp} = \frac{\rho^{\outp} ( X_k^{\outp} - X_k^{\inp})}{\Delta t /2} \enskip,
\end{equation}
and the nuclear energy generation rate is defined as
\begin{equation}
(\rho\Hnuc)^{\outp} = -\sum_k q_k (\rho\omegadot_k)^{\outp} \enskip,
\end{equation}
for the case where only strong-mediated reactions are involved.  As
$\Hnuc$ is an output of the reaction network, more general networks
(i.e., involving weak interactions) can easily be added.  The enthalpy
update incorporates the external heating, $(\rho\Hext)^{\inp}$, and is
updated by
\begin{equation}
(\rho h)^{\outp} = 
(\rho h)^{\inp} + \frac{\dt}{2} (\rho\Hnuc)^{\outp} + \frac{\dt}{2} (\rho\Hext)^{\inp}\enskip.
\end{equation}

\item {\bf Advect Base Density}$[\rho_0^\inp,w_0^\inp]$
$\rightarrow [\rho_0^\outp,\rho_0^{\outp,n+\myhalf}]$\\
is the process by which we update the base state density through $\dt$ in time
as well as compute time-centered interface states, $\rho_0^{\outp,n+\myhalf}$, 
for use later in the algorithm.
\begin{description}

\item[plane-parallel:] We discretize the base state density evolution 
equation from paper III equation (29) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are computed using the
procedure described in paper II, Appendix A.\\

\item[spherical:] The base state density update now includes the area factors in the 
divergences:
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
          \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right] \enskip. 
%\item {\bf Advect Base Pressure}
%is the process by which we update the base state pressure through $\dt$ in time and
%compute $\psi$.  The input arguments are different for plane-parallel and spherical.
%\begin{description}
%\item {\bf Advect Base Pressure Planar}$[p_0^\inp,w_0^\inp,\etarho^\inp] \rightarrow 
%[p_0^\outp,\psi^\outp]$:
%First we compute a cell-centered $\psi^\outp$ using
%\begin{equation}
%\psi_j^\outp = 
%\frac{1}{2} \left(\eta_{\rho,j-\myhalf}^\inp + \eta_{\rho,j+\myhalf}^\inp\right) g \enskip .
%\end{equation}
%Then, we discretize the base state pressure evolution equation from paper III equation (31),
%\begin{equation}
%\frac{\partial p_0}{\partial t} = 
%-w_0 \frac{\partial p_0}{\partial r} + \psi \enskip , \label{eq:p0upd_new}
%\end{equation}
%to compute the new base state pressure,
%\begin{equation}
%p_{0,j}^\outp = p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
%+ w_{0,j-\myhalf}^\inp \right)
%\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) 
%+ \dt \; \psi_j^\outp \enskip,
%\end{equation}
%where the interface states, $p_0^{n+\myhalf}$, are again found via the procedure 
%described in paper II, Appendix A.\\
%\item {\bf Advect Base Pressure Spherical}$[p_0^{\inp},w_0^{\inp},\gammabar^{\inp},
%\Sbar^{\inp}, \rho^\inp, X_k^\inp] \rightarrow [p_0^{\outp},\psi^{n,\outp}, 
%\psi^{n+\myhalf,\outp}]$:
%First use a predictor-corrector formulation for the pressure update in order to improve 
%the overall accuracy.  Beginning with equation (\ref{eq:w01}), we can write
%\begin{equation}
%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} = 
% \Sbar - \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) 
%- \frac{1}{\gammabar p_0} w_0 \frac{\partial p_0}{\partial r} \enskip .
%\end{equation}
%We time-center the left hand side:
%\begin{equation}
%\frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \approx
%\frac{2}{\gammabar^{\inp} (p_0^{\inp} + p_0^{\star})} \frac{p_0^{\star} 
%- p_0^{\inp}}{\dt}\enskip.
%\end{equation}
%We can then solve for the provisional updated pressure, $p_0^{\star}$, giving:
%\begin{equation}
%p_0^{\star} = p_0^{\inp} 
%\frac{\left(1 + \gammabar^{\inp} f\right)}{\left(1 - \gammabar^{\inp} f\right)}\enskip,
%\end{equation}
%with:
%\begin{equation}
%f = \frac{\dt}{2} \left [ \Sbar^{\inp} - 
%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
%\frac{1}{\gammabar^{\inp} p_0^{\inp}} w_0^{\inp} 
%\left ( \frac{\partial p_0^\inp}{\partial r} \right ) \right ] \label{eq:frhs} \enskip ,
%\end{equation}
%where the $(\partial p_0^\inp / \partial r)$ is computed by interpolating $p_0^\inp$ from 
%cell-centers to edges and then differencing.  Next, we compute a provisional updated 
%$\gammabar$ using the provisional updated pressure,
%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^\star,X_k^{\inp})]$,
%where $\rho^\inp$ and $X_k^\inp$ are time-advanced versions of the $\rho$ and
%$X_k$ which were used to compute $\gammabar^\inp$.\\
%Now we begin the corrector step.  Using $p_0^\star$ and $\gammabar^\star$, 
%we can compute a more accurate approximation to equation (\ref{eq:frhs}),
%\begin{equation}
%f^\dagger = \frac{\dt}{2} \left [ \Sbar^{\inp} -
%\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0^{\inp} \right) - 
%\frac{1}{\gammabar^{\dagger} p_0^\dagger} w_0^{\inp} 
%\left ( \frac{\partial p_0^\dagger}{\partial r} \right ) \right ] \enskip ,
%\end{equation}
%where $p_0^\dagger = (p_0^\inp + p_0^\star)/2$ and
%$\gammabar^\dagger = (\gammabar^\inp + \gammabar^\star)/2$.
%This allows us to complete the corrector for the pressure update:
%\begin{equation}
%p_0^{\outp} = p_0^{\inp} \frac{\left(1 + \gammabar^{\dagger} f^\dagger\right)}{\left(1 - \gammabar^{\dagger} f^\dagger\right)}\enskip .
%\end{equation}
%Next, we compute $\psi^{\outp,n}$ and $\psi^{\outp,n+\myhalf}$.  First, we compute an 
%updated $\gammabar$ using the update pressure, 
%$\gammabar^\star = {\bf Avg}[\Gamma_1(\rho^{\inp},p_0^{\outp},X_k^{\inp})]$.  Then, 
%rewriting equation (\ref{eq:w01}), we can define
%\begin{eqnarray}
%\psi \equiv \frac{D_0 p_0}{Dt} &=& \gammabar p_0 \left [ \overline{S} - 
%       \frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 w_0 \right ) \right] \enskip .
%\end{eqnarray}
%We discretize this locally as:
%\begin{equation}
%\psi_j^{\outp,n} = 
%\left(\gammabar^\inp p_0^\inp\right)_j
%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip ,
%\end{equation}
%\begin{equation}
%\psi_j^{\outp,n+\myhalf} = 
%\left(\frac{\gammabar^\inp+\gammabar^{\star}}{2}\right)_j
%\left(\frac{p_0^\inp+p_0^\outp}{2}\right)_j
%\left \{ \Sbar_j^{\inp} -  \frac{1}{r_j^2} \left [ (r^2 w_0^{\inp})_{j+\myhalf} -
%          (r^2 w_0^{\inp})_{j-\myhalf} \right ] \right \} \enskip .
%\end{equation}
%\end{description}
\item {\bf Enforce HSE}$[p_0^{\inp},\rho_0^{\inp},g^{\inp}] \rightarrow [p_0^{\outp}]$
has replaced {\bf Advect Base Pressure} from paper III as the process by which we 
update the base state pressure.  Rather than discretize the evolution equation
for $p_0$, we enforce hydrostatic equilibrium directly, which is simpler and
analytically equivalent.  We update $p_0$ using:
p_{0,j}^{\outp} = p_{0,j-1}^{\outp} + 
\frac{\Delta r}{2}\left(\rho_{0,j}^{inp}+\rho_{0,j-1}^{\inp}\right)g_{j-\myhalf}^{\inp}\enskip .
Once the density is less than the cutoff density, we set $p_{0,j}^\outp = p_{0,j-1}^\outp$.
We use the boundary condition, $p_{0,jmax}^\outp = p_{0,jmax}^\inp$.

@@

We discretize the base state enthalpy equation,
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states $(\rho h)_0^{n+\myhalf}$, are again found via the procedure 
described in paper II, Appendix A.\\

@@

The base state enthalpy update now includes the area factors in the divergences:
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left \{ \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j+\myhalf} -  
          \left[ r^2 (\rho h)_0^{\outp} w_0^{\inp}\right]_{j-\myhalf} \right\}
+\Delta t\psi^{\inp,n+\myhalf} \enskip. 
\end{equation}
To compute the interface states, we use $\psi^{\inp,n}$ as the forcing term, whereas in the
enthalpy update, we use $\psi^{\inp,n+\myhalf}$ as the forcing.

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to reset the base state enthalpy to be
the average of the full state enthalpy since we do not need to compute the complete
evolution of the base state enthalpy.
Unlike paper III where we used $\nabla\cdot\etarho$ as an adjustment to $\rho_0$, 
we instead simply set $\rho_0=\overline{\rho}$, which is analytically equivalent.
\begin{description}
\item[plane-parallel:] Set $\rho_0^{\outp} = {\bf Avg}(\rho^{\inp})$.
\item[spherical:] Set $\rho_0^{\outp} = \rho_0^{\inp} + {\bf Avg}(\rho^{\inp} - \rho_0^{\inp})$.
\end{description}

@@

\item {\bf Compute Gravity}$[\rhozero^{\inp}] \rightarrow [g^{\outp}]$\\
is the process by which we compute gravity at both radial cell-centers and edges.
\begin{description}
\item For plane-parallel, gravity constant in space and time.
\item For spherical, we define the enclosed mass at edges, and then cell-centers:
\begin{equation}
m_{j-\myhalf} = \sum_{k=1}^j \frac{4}{3}\pi\left(r_{k-\myhalf}^3 - r_{k-\sfrac{3}{2}}^3\right)\rho_{0,k-1}^{\inp}; ~~~~~ m_j = m_{j-\myhalf} + \frac{4}{3}\pi \left(r_j^3 - r_{j-\myhalf}^3\right)\rho_{0,j}^{\inp} \enskip .
\end{equation}
Then, we define the gravity at both edges and cell-centers:
\begin{equation}
g_{j-\myhalf}^\outp = \frac{Gm_{j-\myhalf}}{r_{j-\myhalf}^2}; ~~~~~ 
g_j^\outp = \frac{Gm_j}{r_j^2}\enskip .
\end{equation}
\end{description}

@@

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3.] {\em React the full state through the first time interval of $\dt / 2.$}

@@

\item {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n]
                   \rightarrow$ $[ \rho^{(1)}, (\rho h)^{(1)}, X_k^{(1)}, T^{(1)},
                                  (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)} ].$

@@

%{\bf Advect Base Pressure Planar}$[p_0^n, w_0^{\nph,\star}, \etarho^{n+\myhalf,\star}]
%\rightarrow [p_0^{n+1,\star},\psi^{\nph,\star}]$.
%\item For spherical,\\
%{\bf Advect Base Pressure Spherical}$[p_0^n,w_0^{\nph,\star},\gammabar^{(1)},
%\overline{S^{\nph,\star}}, \rho^{(2),\star}, X_k^{(2),\star}]$\\
%$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\rightarrow [p_0^{n+1,\star},\psi^{n,\star},\psi^{\nph,\star}]$.
\item {\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star},g^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
